---
title: "swmmr"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
Sys.setenv(LANG = "en")
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/swmmr)](https://cran.r-project.org/package=swmmr) [![Build Status](https://travis-ci.org/dleutnant/swmmr.svg?branch=dev)](https://travis-ci.org/dleutnant/swmmr)

Functions to connect the widely used [Storm Water Management Model (SWMM)](https://www.epa.gov/water-research/storm-water-management-model-swmm)
of the United States Environmental Protection Agency (US EPA) to R with currently 
two main goals: (1) Run a SWMM simulation from R and (2) provide fast 
access to simulation results, i.e. SWMM's binary '.out'-files. High performance is 
achieved with help of Rcpp. Additionally, reading SWMM's '.inp' and '.rpt' files is supported to 
glance model structures and to get direct access to simulation summaries.

## Installation

Installation is easy thanks to CRAN:
```{r cran, eval = FALSE}
install.packages("swmmr")
```

You can install the dev version from github with:
```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("dleutnant/swmmr@dev")
```

## Example

This is a basic example which shows you how to work with the package. We use 
the example shipped with the SWMM5 executable.

### Initiate a SWMM run and retrieve simulation results

```{r example}

library(swmmr)
library(ggplot2)
library(purrr) # to conveniently work with list objects

# set path to inp (avoid white spaces in file paths!)
inp_path <- "~/EPA_SWMM_Projects/Examples/Example1.inp"

# glance model structure, the result is a list of data.frames with SWMM sections
inp <- read_inp(x = inp_path)

# available SWMM sections
summary(inp)

# for example, inspect section subcatchments
inp$subcatchments

# run a simulation
# the result is a named list of paths, directing
# to the inp, rpt and out-file, respectively.
files <- run_swmm(inp = inp_path)

# we can now read model results from the binary output:
# here, we focus on the system variable (iType = 3) from which we pull
# total rainfall (in/hr or mm/hr) and total runoff (flow units) (vIndex = c(1,4)).
results <- read_out(files$out, iType = 3, vIndex = c(1, 4))

# results is a list object containing two time series 
str(results, max.level = 2)

# basic summary
results[[1]] %>% purrr::map(summary)

# basic plotting
results[[1]] %>% purrr::imap( ~ plot(.x, main = .y))

# We also might be interested in the report file:
# use read_rpt to get is a list of data.frames with SWMM summary sections
report <- read_rpt(files$rpt)

# glance available summaries
summary(report)

# convenient access to summaries through list structure
report$subcatchment_runoff_summary
```

### Visualize the model structure

```{r visualization}
# With help of packages 'ggplot2' and 'sf' we can easily plot swmm objects:
# There is a default plot function which provides a fast way to visualize your 
# model, i.e. it plots subcatchments, junctions, links and raingages
library(ggplot2) # (>= 2.2.1.9000)
library(tidyverse)
library(sf)

plot(inp)

# Because of the underlying ggplot structure, we can individually plot and highlight 
# swmm objects:
# convert subcatchments to sf geometry
sub_sf <- subcatchments_to_sf(inp)
# convert links
lin_sf <- links_to_sf(inp)
# convert junctions
jun_sf <- junctions_to_sf(inp)
# convert raingages
rg_sf <- raingages_to_sf(inp)

# calculate coordinates (centroid of subcatchment) for label position
lab_coord <- sub_sf %>% 
  sf::st_centroid() %>%
  sf::st_coordinates() %>% 
  tibble::as_tibble()

# raingage label
lab_rg_coord <- rg_sf %>% 
  {sf::st_coordinates(.) + 500} %>% # add offset
  tibble::as_tibble()
  
# add coordinates to sf tbl
sub_sf <- dplyr::bind_cols(sub_sf, lab_coord)
rg_sf <- dplyr::bind_cols(rg_sf, lab_rg_coord)

# create the plot
ggplot() + 
  # first plot the subcatchment and colour continously by Area
  geom_sf(data = sub_sf, aes(fill = Area)) + 
  # label by subcatchments by name
  geom_label(data = sub_sf, aes(X, Y, label = Name), alpha = 0.5, size = 3) +
  # add links and highlight Geom1
  geom_sf(data = lin_sf, aes(colour = Geom1), size = 2) +
  # add junctions
  geom_sf(data = jun_sf, aes(size = Elevation), colour = "darkgrey") + 
  # finally show location of raingage
  geom_sf(data = rg_sf, shape = 10) + 
  # label raingage
  geom_label(data = rg_sf, aes(X, Y, label = Name), alpha = 0.5, size = 3) +
  # change scales
  scale_fill_viridis_c() +
  scale_colour_viridis_c(direction = -1) +
  # add labels
  labs(title = "Visualization of a SWMM model", 
       subtitle = "with help of swmmr, ggplot2 and sf")

```

## Contributions
This Git repository uses the [Git Flow](http://nvie.com/posts/a-successful-git-branching-model/) branching model (the [`git flow`](https://github.com/petervanderdoes/gitflow-avh) extension is useful for this). The [`dev`](https://github.com/dleutnant/swmmr/tree/dev) branch contains the latest contributions and other code that will appear in the next release, and the [`master`](https://github.com/dleutnant/swmmr) branch contains the code of the latest release, which is exactly what is currently on [CRAN](https://cran.r-project.org/package=swmmr).

Contributing to this package is easy. Just send a [pull request](https://help.github.com/articles/using-pull-requests/). When you send your PR, make sure `dev` is the destination branch on the [swmmr repository](https://github.com/dleutnant/swmmr). Your PR should pass `R CMD check --as-cran`, which will also be checked by <a href="https://travis-ci.org/dleutnant/swmmr">Travis CI</a> when the PR is submitted.

## Code of condcut
Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Acknowledgments
This package has been mainly developed in the course of the project [STBMOD](https://www.fh-muenster.de/forschung/forschungskatalog/projekt.php?pr_id=722), 
carried out at the [Institute for Infrastructure, Water, Resources, Environment (IWARU)](https://en.fh-muenster.de/iwaru/index.php) of the
[Muenster University of Applied Sciences](https://www.fh-muenster.de).
The project was funded by the German Federal Ministry of Education and Research (BMBF, FKZ 03FH033PX2).

The development of the R package was inspired by the work of [Peter Steinberg](https://github.com/PeterDSteinberg/RSWMM). Also, it benefits from the Interface Guide of [SWMM](https://www.epa.gov/water-research/storm-water-management-model-swmm).

## Citation

```{r citation, echo=FALSE, results='asis'}
citation("swmmr")
```