---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# swmmr

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/swmmr)](https://cran.r-project.org/package=swmmr) [![Build Status](https://travis-ci.org/dleutnant/swmmr.svg?branch=dev)](https://travis-ci.org/dleutnant/swmmr)

Functions to connect the widely used [Storm Water Management Model (SWMM)](https://www.epa.gov/water-research/storm-water-management-model-swmm)
of the United States Environmental Protection Agency (US EPA) to R with currently 
two main goals: (1) Run a SWMM simulation from R and (2) provide fast 
access to simulation results, i.e. SWMM's binary '.out'-files. High performance is 
achieved with help of Rcpp. Additionally, reading SWMM's '.inp'-files is supported to 
glance model structures. 

## Installation

Installation is easy thanks to CRAN:
```{r cran, eval = FALSE}
install.packages("swmmr")
```

You can install the dev version from github with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("dleutnant/swmmr@dev")
```

## Example

This is a basic example which shows you how to work with the package. We use 
the example shipped with the SWMM5 executable.

```{r example}

library(swmmr)
library(purrr) # to conveniently work with list objects

# set path to inp (avoid white spaces in file paths!)
inp_path <- "~/EPA_SWMM_Projects/Examples/Example1.inp"

# glance model structure, the result is a list of data.frames with SWMM sections
inp <- read_inp(inp = inp_path)

# available SWMM sections
summary(inp)

# for example, inspect section timeseries
inp$timeseries

# run a simulation
# the result is a named list of paths, directing
# to the inp, rpt and out-file, respectively.
files <- run_swmm(inp = inp_path)

# we can now read model results from the binary output:
# here, we focus on the system variable (iType = 3) from which we pull
# total rainfall (in/hr or mm/hr) and total runoff (flow units) (vIndex = c(1,4)).
results <- read_out(files$out, iType = 3, vIndex = c(1, 4))

# results is a list object containing two time series 
str(results, max.level = 2)

# basic summary
results[[1]] %>% purrr::map(summary)

# basic plotting
results[[1]] %>% purrr::imap( ~ plot(.x, main = .y))

```

## Contributions
This Git repository uses the [Git Flow](http://nvie.com/posts/a-successful-git-branching-model/) branching model (the [`git flow`](https://github.com/petervanderdoes/gitflow-avh) extension is useful for this). The [`dev`](https://github.com/dleutnant/swmmr/tree/dev) branch contains the latest contributions and other code that will appear in the next release, and the [`master`](https://github.com/dleutnant/swmmr) branch contains the code of the latest release, which is exactly what is currently on [CRAN](https://cran.r-project.org/package=swmmr).

Contributing to this package is easy. Just send a [pull request](https://help.github.com/articles/using-pull-requests/). When you send your PR, make sure `dev` is the destination branch on the [swmmr repository](https://github.com/dleutnant/swmmr). Your PR should pass `R CMD check --as-cran`, which will also be checked by <a href="https://travis-ci.org/dleutnant/swmmr">Travis CI</a> when the PR is submitted.

## Code of condcut
Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Acknowledgments
This package has been developed in the course of the project [STBMOD](https://www.fh-muenster.de/forschung/forschungskatalog/projekt.php?pr_id=722), 
carried out at the [Institute for Infrastructure, Water, Resources, Environment (IWARU)](https://en.fh-muenster.de/iwaru/index.php) of the
[Muenster University of Applied Sciences](https://www.fh-muenster.de).
The project was funded by the German Federal Ministry of Education and Research (BMBF, FKZ 03FH033PX2).

The development of the R package was inspired by the work of [Peter Steinberg](https://github.com/PeterDSteinberg/RSWMM). Also, it benefits from Interface Guide of SWMM. 